---
title: spliceQTL
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This vignette requires both a local machine and the BBMRI virtual machine. On the virtual machine, execute this chunk to set the library path, update the R package spliceQTL, and set the working directory.

```{r virtual machine,eval=FALSE}
lib <- "/virdir/Scratch/arauschenberger/library"
.libPaths(lib)
devtools::install_github("rauschenberger/spliceQTL",lib=lib)
library("spliceQTL",lib.loc=lib)
path <- "/virdir/Scratch/arauschenberger/spliceQTL"
setwd(path)

#R
#rnorm(10)
#Sys.sleep(2)
#q()
#n
```

On a local machine with PLINK, execute this chunk to obtain the Geuvadis SNP data. Then move the files from the local to the virtual machine.

```{r obtain (local),eval=FALSE}
for(chr in 1:22){ 
    spliceQTL::get.snps.geuvadis(chr=chr,data="N:/semisup/data/eQTLs",
                                 path="N:/spliceQTL/data/Geuvadis")
}
```

On the virtual machine, execute this chunk to obtain the BBMRI SNP data, the Geuvadis exon data, and the BBMRI exon data. Choose one out of the six biobanks (CODAM, LL, LLS, NTR, PAN, RS).

```{r obtain (remote),eval=FALSE}
for(chr in 1:22){
    spliceQTL::get.snps.bbmri(chr=chr,biobank="LLS",path=path,size=500*10^3)
}
spliceQTL::get.exons.geuvadis(path=path)
spliceQTL::get.exons.bbmri(path=path)
```

On the virtual machine, execute this chunk to prepare the data. See the documentation of the R package spliceQTL for further information.

```{r prepare,eval=FALSE}
for(chr in 1:22){
  for(data in c("Geuvadis","LLS")){
      
    #wait <- TRUE
    #while(wait){
    #    wait <- !file.exists(paste0(data,".chr",chr,".RData"))
    #    if(wait){Sys.sleep(60);cat(".")}
    #}
      
    rm(list=setdiff(ls(),c("data","chr","path"))); gc()
    set.seed(1)
    
    cat("Analysing",data,chr,":",as.character(Sys.time()),"\n")
    if(data=="Geuvadis"){
      load(file.path(path,"Geuvadis.exons.RData"),verbose=TRUE)
    } else {
      load(file.path(path,"BBMRI.exons.RData"),verbose=TRUE)
      cond <- sapply(strsplit(x=rownames(exons),split=":"),function(x) x[[1]]==data)
      exons <- exons[cond,]
    }
    load(file.path(path,paste0(data,".chr",chr,".RData")),verbose=TRUE)
    
    cat("Matching samples:","\n")
    list <- spliceQTL::match.samples(exons,snps)
    exons <- list$exons; snps <- list$snps; rm(list)
    
    cat("Adjusting samples:","\n")
    exons <- spliceQTL::adjust.samples(x=exons) # slow!
    exons <- asinh(x=exons)
    
    cat("Adjusting covariates:","\n")
    names <- strsplit(x=colnames(exons),split="_") # exon names
    length <- sapply(names,function(x) as.integer(x[[3]])-as.integer(x[[2]])) # exon length
    exons <- spliceQTL::adjust.covariates(x=exons,group=gene_id,offset=length) # slow!
    
    # subset chromosome
    cond <- sapply(strsplit(x=colnames(exons),split="_"),function(x) x[[1]]==chr)
    exons <- exons[,cond]
    gene_id = gene_id[cond] # correct?
    
    cat("Mapping exons:","\n")
    map <- list()
    map$genes <- spliceQTL::map.genes(chr=chr,path=path)
    map$exons <- spliceQTL::map.exons(gene=as.character(map$genes$gene_id),exon=gene_id)
    
    cat("Mapping SNPs:","\n")
    names <- strsplit(x=colnames(snps),split=":")
    snp.chr <- sapply(names,function(x) as.integer(x[[1]]))
    snp.pos <- sapply(names,function(x) as.integer(x[[2]]))
    map$snps <- spliceQTL::map.snps(gene.chr=map$genes$chr,
                                    gene.start=map$genes$start,
                                    gene.end=map$genes$end,
                                    snp.chr=snp.chr,snp.pos=snp.pos)
    
    cat("Dropping genes:","\n")
    map <- spliceQTL::drop.trivial(map=map)
    
    cat("Testing:",as.character(Sys.time())," -> ")
    rm(list=setdiff(ls(),c("exons","snps","map","data","chr","path"))); gc()
    
    save(list=c("exons","snps","map"),file=file.path(path,paste0("temp.",data,".chr",chr,".RData")))
    
#q()
#n
#exit
  }

}
```

On the virtual machine, first restart R and then execute this chunk to test for alternative splicing. (It seems that lme4::lmer in spliceQTL::adjust.covariates fails to release  memory on Linux.)

```{r test,eval=FALSE}
for(chr in 22:1){
    for(data in c("Geuvadis","LLS")){
        cat("Analysing",data,chr,":",as.character(Sys.time()),"\n")
        
        #wait <- TRUE
        #while(wait){
        #    wait <- !file.exists(paste0("temp.",data,".chr",chr,".RData"))
        #    if(wait){Sys.sleep(60);cat(".")}
        #}  
        
        rm(list=setdiff(ls(),c("data","chr","path"))); gc()
        load(file.path(path,paste0("temp.",data,".chr",chr,".RData")))
        pvalue <- spliceQTL::test.multiple(Y=exons,X=snps,map=map,rho=1,spec=20)
        save(object=pvalue,file=file.path(path,paste0("pval.",data,".chr",chr,".RData")))
    }
#q()
#n
#exit
}
```

On the virtual machine, execute this chunk to compare the results between the Geuvadis and the BBMRI project.

```{r,eval=FALSE}
cor <- chisq <- rep(NA,length=22)
for(chr in 22:1){
    load(file.path(path,paste0("pval.Geuvadis.chr",chr,".RData")))
    a <- pvalue; pvalue <- NA
    load(file.path(path,paste0("pval.LLS.chr",chr,".RData")))
    b <- pvalue; pvalue <- NA

    names <- intersect(rownames(a),rownames(b))
    plot(jitter(-log(a[names,1])),jitter(-log(b[names,1])))
    cor[chr] <- stats::cor(a[names,1],b[names,1],method="spearman")

    a <- stats::p.adjust(a[names,1])<0.05
    b <- stats::p.adjust(b[names,1])<0.05
    print(paste0("chr",chr))
    print(table(a,b))
    chisq[chr] <- stats::chisq.test(table(a,b))$p.value
}
```


memory and CPU usage Linux: htop


